---
title: "Mais do R"
author: "Marcelo Linardi"
date: "2025-03-11"
output: html_document
---

<!-- ## Setup -->

```{r setup, echo=FALSE, include=FALSE}
library(knitr)
# knitr::opts_chunk$set(echo = TRUE)
library(haven)
library(rmarkdown)
library(writexl)
library(Hmisc)
library(naniar)
library(here)
library(tidyverse)
library(readxl)
library(visdat)
library(stargazer)

rm(list = ls())
here()
getwd()


```



## Load your data 
```{r}
nlswork <- as.data.frame(read_excel("nlswork.xlsx"))
# nlswork <- read_dta("nlswork.dta") # in case you have a Stata data source

```

## Define the object as a DataFrame

```{r}
df <- as.data.frame(nlswork)

```


# Data manipulation -- check the pipe operator, %>%


## Select a subset of variables

```{r}
nlswork_s<- nlswork %>% 
  select(idcode, ln_wage) 

```

## Rename variables

```{r}
nlswork_r <- nlswork %>% 
  rename(cae = ind_code)

```

## Filter a subset of observations

```{r}
nlswork_f<- nlswork %>% 
  filter(age > 40) 

```

## Mutate: create variables

```{r}
# nlswork_m <- nlswork %>% 
#   mutate(ln_asd=ln(age))

```

## Manipulate the data in a single sequence

```{r}
nlswork1<- nlswork %>% 
  rename(cae = ind_code) %>%
  select(idcode, ln_wage, age) %>% 
  filter(age > 40) %>%
  mutate(age2=age^2)

```


# Descriptive statistics

```{r}
# install.packages("stargazer")
library(stargazer)
df %>% 
  count()

summary(df)
table(df$race, df$collgrad)
str(df)

```


```{r, echo = FALSE, results='asis'}
    df %>%
      dplyr::select(collgrad,ttl_exp,hours) %>% 
      stargazer(title="Shorter statistics",
      type= "text",
      out = "Stats.html",
      covariate.labels = c("Collage Graduate","Experience","Hours"),
      digits = 2, summary.stat = c("min","mean","sd","max"),
      header = FALSE)
```

## Export descriptive statistics table to html, with 2 digits

```{r}
nlswork %>%
  dplyr::select(age, collgrad, ttl_exp, union, hours) %>% 
  stargazer(title="Shorter statistics",
            type= "text", out = "Statistics_output.html",
            digits = 2)

```

## Export descriptive statistics table to txt, with 3 digits

```{r}
nlswork %>%
  dplyr::select(age, collgrad, ttl_exp, union, hours) %>% 
  stargazer(title="Shorter statistics",
            type= "text", out = "Statistics_output.txt",
            digits = 3)

```

## Transposing the descriptive statistics table

```{r}
nlswork %>%
  dplyr::select(age, collgrad, ttl_exp, union, hours) %>% 
  stargazer(title="Shorter statistics",
            type= "text", out = "Statistics_output2.txt",
            digits = 3, flip=TRUE)

```



# Visualize missing information:

```{r}
# install.packages("visdat")
# install.packages("naniar")
library(naniar)
vis_miss(nlswork)

```

```{r}
gg_miss_var(nlswork) + labs(y = "Total missing values for each variable")
gg_miss_upset(nlswork)

```

```{r}
n_var_miss(nlswork)

ggplot(nlswork,aes(x=age,y=ln_wage))+
  geom_point()

ggplot(nlswork,aes(x=age,y=ln_wage))+
  geom_miss_point()

ggplot(nlswork,aes(x=age,y=ln_wage))+
  geom_miss_point() +
  facet_wrap(~race)

gg_miss_fct(x = nlswork,fct = year)

```


### Alternative

```{r}
library(visdat)
vis_dat(nlswork)

```

# Visualisation to explore your data

## Categorical variable

### My histogram

```{r echo=FALSE, warning=FALSE, results='asis'}
df %>% 
      ggplot(aes(ttl_exp,ln_wage)) +
      labs(title = "Ln Wage vs. Experience") +
      ylab("Ln Wage") +
      xlab("Experience") +
      geom_point()

```

```{r}
ggplot(data = nlswork) +
  geom_bar(mapping=aes(x=as.factor(collgrad))) +
  xlab("College graduate (1=Yes)")

nlswork %>% 
  count(collgrad)
```

  
## Continuous variable

```{r}
ggplot(data = nlswork) +
  geom_histogram(mapping = aes(x = wks_work), binwidth = 0.5)

```

## Categorical and continuous variables

```{r}
df %>% ggplot(aes(x=as.factor(collgrad), y=ln_wage)) +
  geom_boxplot(fill="slateblue", alpha=0.2) + 
  xlab("College graduate (1=Yes)")
```

```{r,warning=FALSE}
df %>% ggplot(mapping = aes(x = ln_wage, y = ..density..)) +
    xlab("ln(wage)") +
    ylab("Density") +
    geom_freqpoly(mapping = aes(colour = factor(collgrad, labels=c("No", "Yes")))) +
  labs(color ="College degree")
```